### 2022-08-20 surv.Rmd
initial

---
title: "Untitled"
author: "Dean Marchiori"
date: "8/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(survival)
library(survminer)
library(SMPD4)
library(dplyr)
```

```{r}
nedmaba <- smpd4_phenotype |> 
  select(id, survival_time, deceased, gender) |> 
  mutate(deceased = as.numeric(deceased)) 
```

```{r}
km1 <- survfit(Surv(survival_time, deceased) ~ 1, data = nedmaba)
summary(km1, times= (0:13)*365, scale=365)
ggsurvplot(km1,
          conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF"))
```

```{r}
km2 <- survfit(Surv(survival_time, deceased) ~ gender, data = nedmaba)
km2
summary(km2, times= (0:13)*365, scale=365)

ggsurvplot(km2,
          pval = TRUE, conf.int = TRUE,
          risk.table = TRUE, # Add risk table
          risk.table.col = "strata", # Change risk table color by groups
          linetype = "strata", # Change line type by groups
          surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("#E7B800", "#2E9FDF"))

survdiff(Surv(survival_time, deceased) ~ gender, data = nedmaba)

fit <- coxph(Surv(survival_time, deceased) ~ gender, data = nedmaba)
summary(fit)
```